Laws for Creating Trust in the Blockchain Age

### 2017-07-25 tex/article.tex
Adjusted footcites, they are now after the period

For these platforms, trust is an essential element: people step into the car of a person they never met before or let strangers sleep in their spare room without worries.
Trust is addressed by rating systems, the foundation for their quality and safety measures.
For instance, pre-testing for potentially dangerous drivers and pre-screening is reduced and replaced with a continuous quality monitoring system\footcite{uberfivestar}.
The Uber mobile application asks both drivers and passengers to provide feedback on their experience.
The software averages the one to five-star ratings of the most recent 500 trips\footcite{uberratinghelp}.
If this average long-term quality rating of a driver drops below a certain level, it automatically triggers contract termination\footcite{forbesuserfiring}. 
The traditional dynamics of trust are altered on such platforms.
Employment regulations are under pressure: some courts found that Uber violated local taxi regulations\footcite{ftuberregulation}\footcite{reutersuberdanish}\footcite{reutersubersouthkorea}.
For the United Kingdom these platforms are responsible for lowering unemployment levels, reducing fixed employment contracts by a few percent, enforce the trend towards zero-hour contracts and self-employment.
A 2017 United Kingdom government report states that currently 1.1 million people work within this gig economy\footcite{goodgigs}.
The laws which govern trust, employment terminations, and access to marketplaces are no longer the exclusive domain of national law.
Laws of trust are now partly inside commercial software.

@@

Blockchain technology is at the core of our proposed Uber alternative and blueprint for a public gig economy.
However, Bitcoin and Ethereum are in our opinion unsuitable for this task or other legal processes such as land registration.
We do not consider cybercurrency research in the remainder of this work.
Our reason is that all current work based on coin creation, blockchain consensus like proof-of-work or variants thereof lack robust and effective ties to any legal system, or even the real world in general.
The current generation of cybercurrency-based blockchain technology lacks a durable governance structure.
Severe disagreements exists within these communities, as various factions battle for control of ecosystems worth billions.
The cybercurrency ecosystem is unsuitable for land registration as it occasionally has periods dubbed civil war by the media\footcite{bloombergbitcoinsplinter}.
We also do not consider private or consortium blockchains, as they do not offer any significant advantage over classical distributed databases.
We focus on the tamper-proof leaderless database type of solutions.
The statements and opinions in this paper are based on our decade-long experience of deploying and gradually improving our own ledger, installed by 1.8 million users.
We deployed a very primitive fully distributed ledger in August 2007, pre-dating the launch of Bitcoin\footcite{bbcnewstribler}.
In 2017 we have mathematically proven that our blockchain technology, TrustChain, scales linear and surpasses a transaction throughput of 10.000 transactions per second\footcite{githubissueconsensus}.

@@

Contracts, transactions, and the records of them are among the defining structures in our economic, legal, and political systems\footcite{harvardbusinessblockchain}.
They protect assets and set organizational boundaries.
These authentic records create ownership, determine citizenship, and define partnerships between entities.
It establishes a governance layer in which states, economic actors, and citizens interact, effectively creating trust.

For thousands of years trusted guardians kept authentic ownership records of land and assets in general.
These centralized bureaucracies which govern these records have not changed much, while the world is slowly being filled with autonomous cars, trucks, drones, and barges. 
This difference in innovation speed is creating increasing levels of friction around data governance. 
The legal world might be at the beginning of a transition towards full digitization, additional standardization, and further automation. 
Already 22 years ago the book "Law in a Digital World" was published, however deeds of land transfer are rarely pure in digital form\footcite{katsh1995law}.

Blockchain enthusiasts claim this technology offers an alternative to any trusted guardian and central bank.
It offers us another way to organise society, with a high degree of decentralization.
The resilience of Bitcoin together with the rising gig economy endorses this claim.
The breakthrough that made the gig economy possible was the use of reputation systems to build trust between strangers.
Reputation systems collect and process information about past interactions, to help people evaluate the trustworthiness of others\footcite{resnick2002trust}.
However, an individual’s reputation on a platform such as eBay is owned by a profit-driven entity.
This leads to the following problems for the social good:
\begin{itemize}
	\item Lock-in: a solid reputation on one platform is locked into that platform: you cannot move your reputation. Like other forms of lock-in, this inhibits competition and encourages monopoly behaviour. Ironically, companies in the \enquote{sharing economy} do not share reputations. Users are increasingly being protected by European Union law from such data silos. The newly defined right to data portability allows individuals to obtain and reuse their personal data for their own purposes across different services\footcite{dataportabilityeu}.
	\item Fragmentation: each company operates its own closed market, only accessible through their (mobile) application. This leads to lower overall efficiency, compared to a single open market. For example, the ride-hailing market is fragmented into numerous closed markets operated by companies such as Uber, Lyft, BlaBla Car, Didi Kuaidi, GrabTaxi and Karhoo. Each isolated marketplace tries to match drivers and customers in real-time and have to overcome similar challenges.
\end{itemize}
Centralized approaches are ultimately always influenced by the goals and the reliability of the central entity or authority that controls them. 
One economic model even predicts the rise of predatory, monopolistic platforms within two-sided markets\footcite{loertscher2016predatory}.
For this reason, central platforms can never be truly generic and universal. 
With a decentralized approach, multiple independent individuals cooperate in some manner to build a single ecosystem and no one entity has control over the entire environment.
This is the model we envision for the future programmable economy: trust relations are not locked to a single profit-driven entity.

@@

In May 1962 the vision of a time-sharing computer system with many remote stations was presented, known as the Internet today\footcite{licklider1962line}.
Nobody owns the Internet.
This has been a critical factor for its decades long success story.
The Internet consists of numerous Autonomous Systems which are loosely coupled and have a common numbering mechanism.
On top of this global communication infrastructure we built email, video conferencing, entertainment platforms, search engines, marketplaces, countless cloud services, and essentially a digital economy.
However, these examples often contain a single central point of control and authority.
It has proven to be hard to create a decentralized governance layer for such vital public infrastructure.

Joseph Stiglitz co-authored the "Architecture of Economic Systems" in 1985, describing how decision making units can be organized together within a system\footcite{sah1985architecture}.
It presents a basic framework to compare the performance of decentralized and centralized economic systems.
Centralized economic systems have given way to decentralized forms.
The current challenge is to further refine the decentralized form: storage and governance of authoritative answers to ownership questions using a public and transparent infrastructure.

Trust, accountability, and reputation mechanisms are closely tied.
Transparency can be used to make an authority accountable in order to establish trust. 
It promotes integrity of operations by monitoring the correct behaviour of economic actors\footcite{troncoso2017systematizing}.

@@

We believe that blockchain technology enables decentralized economies of higher efficiency and stability then currently known. 
This work broadens the applications of the organisational principle behind the Internet, offering full sovereignty to all decision making entities, full transparency on their past performance, and openness in general.
We are applying ideas fundamentally concerned with freedom of individuals, embedding humanities natural disposition to trust, and combining it with a very unforgiving blockchain-based mechanism for rule-breaking economic actors: digital ostracism\footcite{bicchieri2004trust}. 
The threat of being banned forever from an ecosystem may seem to lack compassion and decency, but boosts trust. 
Game theory shows that the "shadow of the future" removes cheating incentives\footcite{friedman1971non}.
Note that fraud differs from blunders by economic actors, which merely impact their reputation (e.g. star rating).
In our architecture we apply the grim trigger defined in 1971 to punish dishonesty, lying, cheating, and fraud.
In other words, mistakes can be forgiven, but intentional digital manipulations are not.

@@

Figure \ref{fig:tech_stack_simple} shows the four layers which define our architecture.
First we need to address the digital identity problem, shown as the bottom layer.
A wealth of applications require strong authentication and long-lived secure identities.
The Internet requires a common continuously evolving strong identity layer.
This would make the Internet safer, better and more efficient.
A single common identity layer also needs full decentralization and self-governance.
This blocks progress, for instance, a single standard for globally legally valid electronic signatures which does not yet exist.
This work is motivated by the lack of control over our identity, as formulated within the proposed WebDHT documentation\footcite{webdhtproposal}:

@@

Self-sovereign identities require that users are the rulers of their own identity.
This idea was presented in 2012 by Moxie Marlinspike\footcite{sovereignsource}.
With this concept people and businesses can store their own identity data on their own devices, and present it efficiently to anyone who wants to validate it.
The decentralized storage within the concept removes the need for any central database of identity data.
This might be increasingly important with the European Union General Data Protection Regulation\footcite{dataportabilityeu}.

@@

Being the ruler of your own identity is closely related to the core concept of Bitcoin.
Bitcoin creates money without banks, but also provides users with an exceptional level of control.
The novelty of Bitcoin is that is provides a system where nobody can stop a specific entity from spending their money\footcite{r3corda}.
Identity autonomy provides a new level of citizen empowerment and enables financial sovereignty\footcite{matouk2009financial}.

@@

TrustChain and all other layers in our architecture utilise a new model of governance that we call self-governance.
It is an ecosystem where ordinary users self-organise into a large-scale Internet-based collective which can be freely joined by anyone, with the special strong property that all authority is temporary.
We define a self-governance system as a distributed system in which autonomous individuals can collectively exercise all of the necessary functions of power without intervention from any authority which they cannot themselves alter.
Self-governance implies a mechanism for peaceful transfer of power. 
Users may directly vote on new laws, changes to rules, routine maintenance updates, and key changes in principles. 
This approach has proven to be usable, but occasionally leads to fascinating outcomes.
On 20 July 2016 the Ethereum community split apart because a majority voted to revert a completely valid transaction, created by faulty user software (known as the DAO hack)\footcite{cryptocomparedao}.
Another option is to use representatives and appoint professionals responsible for the daily administration of the community through some voting process.
However, this goes against the prevailing anti-authority culture within operational blockchain ecosystems.
Distributed systems with self-governance have no external overseers and no central controlling servers.

The third layer in Figure \ref{fig:tech_stack_simple} transforms our usage of money.
Our proposed approach there is again the opposite of the cybercurrency community.
We believe building a trustworthy and reliable financial infrastructure from scratch is unwise.
By re-engineering and building upon existing bank-based payment systems we provide a realistic, legally compliant, and efficient overlay.
We re-use the existing infrastructure and transform it into loosely coupled autonomous entities.
We provide a proof-of-principle prototype for making international money transfers with sub-second speed, near-zero commissions, instant clearing, and real-time settlement. 
Steps towards a programmable economy necessitate a transformation of the expensive transaction network build in 1970s (SWIFT) using a programming language from the 1950s (COBOL)\footcite{scott2017society}.

@@

In a recent economics publication it is stated that with blockchains, marketplaces can be bootstrapped without the need of traditional trusted intermediaries, lowering the cost of networking\footcite{catalini2016some}.
Furthermore they challenge existing revenue models and incumbents's market power, and opens opportunities for novel approaches to regulation, auctions and the provision of public goods, software, identity and reputation systems.
Little experience, evidence, and knowledge exists for devising this fourth layer.
We are currently conducting trials with various businesses to expand our understanding.
This is a new, developing area of research, however, we believe too few research teams have the resources for the learn-by-doing methodology required to make serious scientific progress.
Building open infrastructures with self-governance is costly (our key funding was the EU FP7 project P2P-Next of 19.500.000 Euro\footcite{p2pnextfunding} and QLectives of 6.900.000 Euro\footcite{qlectivefunding}).

@@

Recently the World Bank indicated that 70\% of the world’s population still lacks access to proper land titling or demarcation services\footcite{ieglandadministration}.
A digital land registration system using blockchain technology requires a re-design of existing procedures for submission and verification of records and claims.
It is most likely to demand changes to legal frameworks.

Obtaining agreement from numerous parties during several of the stages of a property transactions, detecting errors, and to preserve tamper-proof logging is a significant challenge.
Central keepers of records are often entrusted with this critical coordination task.
Blockchains clearly have potential here, offering transparency, reducing overhead, and especially providing a single-source-of-truth.
Various reports support this opinion\footcite{deloitteblockchain}.

@@

A Physical Unclonable Function (PUF) is a device that is easy and cheap to produce but practically infeasible to duplicate, due to minor variations in the manufacturing process of the hardware\footcite{cortez2012modeling}.
Their typical usage can be found in applications that require a high level of security, for instance, self-sovereign identity solutions.
The device offers tamper-proof and safe storage of cryptographic keys, representing ones identity.
Each PUF device contains an unique fingerprint, determined by randomness of embedded components.
A PUF responds to challenges and leads to unique but unpredictable responses, together forming a challenge-response pair.
These challenges are often triggered by pushing a physical button attached to the device.

@@

We present a mobile biometric-based authentication prototype that does not involve any central authority\footcite{hammudoglu2017portable}.
A proof-of-principle mobile application has been developed for Android, capable of matching fingerprints using the built-in device camera.
By only utilizing device-specific components like the embedded storage and camera, no permissioned or specialized, contractual hardware is needed.

@@

Most identity services are offered by a single authority, handing out, revoking and managing identities.
This trend continued when society adopted digital solutions: issuers of digital identities like Internet Assigned Numbers Authority (IANA) and Internet Corporation for Assigned Names and Numbers (ICANN), the organizations responsible for coordination of digital namespaces on the Internet, were and still are single, large organizations that have a large influence on the layout of the Internet landscape\footcite{pathtossi}.
Some identity schemes took a small step beyond a centralized structure and offered hierarchical solutions like certificate authorities (CAs).
However, such systems rely on root authorities which form a single point of failure.

@@

TrustChain blocks are designed to be disseminated and replicated throughout the network.
In particular, this is important when transaction history is used as input for a reputation mechanism which needs the historical behaviour of others (see Section \ref{sec:reputation}).
Additionally, replication of blocks makes the system resistant against network churn where users go on- and offline at a fast rate.
Each user operates on their own bulk storage of blocks, resulting in partial storage of the global ledger.
Collecting information of other users is challenging due to the vulnerability to various attacks, their limited resources and the burst of their interactions.
Prior work investigates this problem and propose solutions for reliable and secure collection of the interaction history\footcite{gkoroureducing}.

@@

\subsection{Scalable Consensus}
\label{sec:scalable_consensus}
Most blockchain fabrics have a requirement for a consistent state which is agreed upon by (a part of) the network.
This requirement holds in particular when considering digital money (cybercurrencies) like Bitcoin or Ethereum.
Consensus of the global ledger is necessary to prevent the double-spending attack where users transfer the same digital asset twice in multiple transactions.
Possibility of such an attack impacts reliability and trust of the overall system, preventing community adoption.
Unfortunately, most consensus mechanisms are computationally expensive and limit the global transaction throughput of the system.
For instance, the proof-of-work consensus mechanism implemented in the Bitcoin fabric limits the theoretical transaction throughput to around seven transactions per second which is by far not enough for a medium to large-sized trading platform\footcite{duebitcoinscale}.
In comparison, Visa processes several hundreds of transactions every second and in April 2017, SWIFT recorded an average of 28.38 million payments per day or around 328 per second\footcite{visatransactions}\footcite{swifttransactions}.
This motivates the need for a scalable blockchain.

@@

Our consensus mechanism proceeds in rounds and each round consists of two phases.
The outcome of each round is a set of checkpoint blocks agreed on by the facilitators of that round.
Facilitators are special user, elected during the first phase in a round of consensus.
These facilitators reach consensus not on the individual transactions like in Bitcoin or Ethereum but on the state of every chain.
If a specific user is not a facilitator, it sends its most recent checkpoint block to all facilitators.
When the facilitators received a sufficient number of checkpoint blocks, they start to reach consensus on all received checkpoint blocks using an Asynchronous Subset Consensus (ACS) algorithm\footcite{miller2016honey}.
When this algorithm is finished, the facilitators send two messages to all users, first the consensus result and second a signature message where the facilitator signed the consensus result, adding authenticity to the consensus result.
When a user receives the consensus result and a sufficient amount of valid facilitator signatures, a new checkpoint block is created and appended to the chain.
Finally, the new set of facilitators is elected and the round number is increased.
The process starts over now.

@@

\subsection{Trustworthiness and Reputation}
\label{sec:reputation}
One way to quantity trustworthiness in an economic system, is by using a reputation mechanism.
Reputation systems are widely used within the gig economy but the trust is often centred to a single authority.
Fifteen years of research on generic distributed reputation systems has yielded a wide variety of proposed designs and possible solutions\footcite{delaviz2010improving}\footcite{delaviz2012sybilres}\footcite{kamvar2003eigentrust}.
Prior work by us challenges the problem of estimating reputation using network flow algorithms.
This mechanism is called BarterCast and has been verified by a real-world implementation in our file-sharing software Tribler\footcite{meulpolder2009bartercast}.
The most challenging attack on decentralized reputation systems is the Sybil attack, where a malicious user manipulates his or her standing in the network by creating multiple fake identities (sybils) and initiating interactions with them.
To date, the Sybil attack remains largely unsolved and solutions that prevent manifestation such an attack, are often complex.

@@

Our reputation mechanism is based on PageRank, designed in 1998 by Larry Page and is called Temporal PageRank since it incorporates the notion of time\footcite{page1999pagerank}\footcite{otte2016sybil}.
PageRank is an accurate model that captures user behaviour when browsing the web and yields a ranking of websites based on relevance for that user.
The algorithm models websites and links between websites as a network and explores this network in a structured manner.
Temporal PageRank works in the same way, exploring the TrustChain data structure and analysing past transactions, including connections between them.
Additionally, it is a relatively simple and cheap technique from a computational perspective, requiring only minimal resources.
Temporal PageRank determines trustworthiness scores for other users from the perspective of a user performing the computation and is somewhat resistant against Sybil attacks in a sense that such attacks performed in the past only have minor influence on the outcome.

@@

To obtain insight in the real-world efficiency of our system, we evaluate matching efficiency of Uber taxi drives.
Uber, one of the largest companies operating in the sharing economy, can be considered as a two-sided market where taxi drivers, offering a ride to a location, are matched against users requesting transportation.
Since there is no public, reliable dataset provided by the Uber platform, we use historical information of taxi rides published by the government of New York instead\footcite{newyorktaxi}.
The dataset provides detailed temporal and geographical information of pick-up and drop-off location of each taxi ride.
We assume a total of 1100 taxi drivers and 1000 users requesting transportation, which is also close to the limitations of the hardware on which the experiment is executed.
It is also assumed that only the taxi drivers perform matching since they are likely to be connected to the network for a longer time whereas passengers most likely close the application when a taxi is available for them.
Taxi drivers and passengers are matched based on their geographical distance.

>>>>

For these platforms, trust is an essential element: people step into the car of a person they never met before or let strangers sleep in their spare room without worries.
Trust is addressed by rating systems, the foundation for their quality and safety measures.
For instance, pre-testing for potentially dangerous drivers and pre-screening is reduced and replaced with a continuous quality monitoring system.\footcite{uberfivestar}
The Uber mobile application asks both drivers and passengers to provide feedback on their experience.
The software averages the one to five-star ratings of the most recent 500 trips.\footcite{uberratinghelp}
If this average long-term quality rating of a driver drops below a certain level, it automatically triggers contract termination.\footcite{forbesuserfiring}
The traditional dynamics of trust are altered on such platforms.
Employment regulations are under pressure: some courts found that Uber violated local taxi regulations.\footcite{ftuberregulation}\footcite{reutersuberdanish}\footcite{reutersubersouthkorea}
For the United Kingdom these platforms are responsible for lowering unemployment levels, reducing fixed employment contracts by a few percent, enforce the trend towards zero-hour contracts and self-employment.
A 2017 United Kingdom government report states that currently 1.1 million people work within this gig economy.\footcite{goodgigs}
The laws which govern trust, employment terminations, and access to marketplaces are no longer the exclusive domain of national law.
Laws of trust are now partly inside commercial software.

@@

Blockchain technology is at the core of our proposed Uber alternative and blueprint for a public gig economy.
However, Bitcoin and Ethereum are in our opinion unsuitable for this task or other legal processes such as land registration.
We do not consider cybercurrency research in the remainder of this work.
Our reason is that all current work based on coin creation, blockchain consensus like proof-of-work or variants thereof lack robust and effective ties to any legal system, or even the real world in general.
The current generation of cybercurrency-based blockchain technology lacks a durable governance structure.
Severe disagreements exists within these communities, as various factions battle for control of ecosystems worth billions.
The cybercurrency ecosystem is unsuitable for land registration as it occasionally has periods dubbed civil war by the media.\footcite{bloombergbitcoinsplinter}
We also do not consider private or consortium blockchains, as they do not offer any significant advantage over classical distributed databases.
We focus on the tamper-proof leaderless database type of solutions.
The statements and opinions in this paper are based on our decade-long experience of deploying and gradually improving our own ledger, installed by 1.8 million users.
We deployed a very primitive fully distributed ledger in August 2007, pre-dating the launch of Bitcoin.\footcite{bbcnewstribler}
In 2017 we have mathematically proven that our blockchain technology, TrustChain, scales linear and surpasses a transaction throughput of 10.000 transactions per second.\footcite{githubissueconsensus}

@@

Contracts, transactions, and the records of them are among the defining structures in our economic, legal, and political systems.\footcite{harvardbusinessblockchain}
They protect assets and set organizational boundaries.
These authentic records create ownership, determine citizenship, and define partnerships between entities.
It establishes a governance layer in which states, economic actors, and citizens interact, effectively creating trust.

For thousands of years trusted guardians kept authentic ownership records of land and assets in general.
These centralized bureaucracies which govern these records have not changed much, while the world is slowly being filled with autonomous cars, trucks, drones, and barges. 
This difference in innovation speed is creating increasing levels of friction around data governance. 
The legal world might be at the beginning of a transition towards full digitization, additional standardization, and further automation. 
Already 22 years ago the book "Law in a Digital World" was published, however deeds of land transfer are rarely pure in digital form.\footcite{katsh1995law}

Blockchain enthusiasts claim this technology offers an alternative to any trusted guardian and central bank.
It offers us another way to organise society, with a high degree of decentralization.
The resilience of Bitcoin together with the rising gig economy endorses this claim.
The breakthrough that made the gig economy possible was the use of reputation systems to build trust between strangers.
Reputation systems collect and process information about past interactions, to help people evaluate the trustworthiness of others.\footcite{resnick2002trust}
However, an individual’s reputation on a platform such as eBay is owned by a profit-driven entity.
This leads to the following problems for the social good:
\begin{itemize}
	\item Lock-in: a solid reputation on one platform is locked into that platform: you cannot move your reputation. Like other forms of lock-in, this inhibits competition and encourages monopoly behaviour. Ironically, companies in the \enquote{sharing economy} do not share reputations. Users are increasingly being protected by European Union law from such data silos. The newly defined right to data portability allows individuals to obtain and reuse their personal data for their own purposes across different services.\footcite{dataportabilityeu}
	\item Fragmentation: each company operates its own closed market, only accessible through their (mobile) application. This leads to lower overall efficiency, compared to a single open market. For example, the ride-hailing market is fragmented into numerous closed markets operated by companies such as Uber, Lyft, BlaBla Car, Didi Kuaidi, GrabTaxi and Karhoo. Each isolated marketplace tries to match drivers and customers in real-time and have to overcome similar challenges.
\end{itemize}
Centralized approaches are ultimately always influenced by the goals and the reliability of the central entity or authority that controls them. 
One economic model even predicts the rise of predatory, monopolistic platforms within two-sided markets.\footcite{loertscher2016predatory}
For this reason, central platforms can never be truly generic and universal. 
With a decentralized approach, multiple independent individuals cooperate in some manner to build a single ecosystem and no one entity has control over the entire environment.
This is the model we envision for the future programmable economy: trust relations are not locked to a single profit-driven entity.

@@

In May 1962 the vision of a time-sharing computer system with many remote stations was presented, known as the Internet today.\footcite{licklider1962line}
Nobody owns the Internet.
This has been a critical factor for its decades long success story.
The Internet consists of numerous Autonomous Systems which are loosely coupled and have a common numbering mechanism.
On top of this global communication infrastructure we built email, video conferencing, entertainment platforms, search engines, marketplaces, countless cloud services, and essentially a digital economy.
However, these examples often contain a single central point of control and authority.
It has proven to be hard to create a decentralized governance layer for such vital public infrastructure.

Joseph Stiglitz co-authored the "Architecture of Economic Systems" in 1985, describing how decision making units can be organized together within a system.\footcite{sah1985architecture}
It presents a basic framework to compare the performance of decentralized and centralized economic systems.
Centralized economic systems have given way to decentralized forms.
The current challenge is to further refine the decentralized form: storage and governance of authoritative answers to ownership questions using a public and transparent infrastructure.

Trust, accountability, and reputation mechanisms are closely tied.
Transparency can be used to make an authority accountable in order to establish trust. 
It promotes integrity of operations by monitoring the correct behaviour of economic actors.\footcite{troncoso2017systematizing}

@@

We believe that blockchain technology enables decentralized economies of higher efficiency and stability then currently known. 
This work broadens the applications of the organisational principle behind the Internet, offering full sovereignty to all decision making entities, full transparency on their past performance, and openness in general.
We are applying ideas fundamentally concerned with freedom of individuals, embedding humanities natural disposition to trust, and combining it with a very unforgiving blockchain-based mechanism for rule-breaking economic actors: digital ostracism.\footcite{bicchieri2004trust}
The threat of being banned forever from an ecosystem may seem to lack compassion and decency, but boosts trust. 
Game theory shows that the "shadow of the future" removes cheating incentives.\footcite{friedman1971non}
Note that fraud differs from blunders by economic actors, which merely impact their reputation (e.g. star rating).
In our architecture we apply the grim trigger defined in 1971 to punish dishonesty, lying, cheating, and fraud.
In other words, mistakes can be forgiven, but intentional digital manipulations are not.

@@

Figure \ref{fig:tech_stack_simple} shows the four layers which define our architecture.
First we need to address the digital identity problem, shown as the bottom layer.
A wealth of applications require strong authentication and long-lived secure identities.
The Internet requires a common continuously evolving strong identity layer.
This would make the Internet safer, better and more efficient.
A single common identity layer also needs full decentralization and self-governance.
This blocks progress, for instance, a single standard for globally legally valid electronic signatures which does not yet exist.
This work is motivated by the lack of control over our identity, as formulated within the proposed WebDHT documentation:\footcite{webdhtproposal}

@@

Self-sovereign identities require that users are the rulers of their own identity.
This idea was presented in 2012 by Moxie Marlinspike.\footcite{sovereignsource}
With this concept people and businesses can store their own identity data on their own devices, and present it efficiently to anyone who wants to validate it.
The decentralized storage within the concept removes the need for any central database of identity data.
This might be increasingly important with the European Union General Data Protection Regulation.\footcite{dataportabilityeu}

@@

Being the ruler of your own identity is closely related to the core concept of Bitcoin.
Bitcoin creates money without banks, but also provides users with an exceptional level of control.
The novelty of Bitcoin is that is provides a system where nobody can stop a specific entity from spending their money.\footcite{r3corda}
Identity autonomy provides a new level of citizen empowerment and enables financial sovereignty.\footcite{matouk2009financial}

@@

TrustChain and all other layers in our architecture utilise a new model of governance that we call self-governance.
It is an ecosystem where ordinary users self-organise into a large-scale Internet-based collective which can be freely joined by anyone, with the special strong property that all authority is temporary.
We define a self-governance system as a distributed system in which autonomous individuals can collectively exercise all of the necessary functions of power without intervention from any authority which they cannot themselves alter.
Self-governance implies a mechanism for peaceful transfer of power. 
Users may directly vote on new laws, changes to rules, routine maintenance updates, and key changes in principles. 
This approach has proven to be usable, but occasionally leads to fascinating outcomes.
On 20 July 2016 the Ethereum community split apart because a majority voted to revert a completely valid transaction, created by faulty user software (known as the DAO hack).\footcite{cryptocomparedao}
Another option is to use representatives and appoint professionals responsible for the daily administration of the community through some voting process.
However, this goes against the prevailing anti-authority culture within operational blockchain ecosystems.
Distributed systems with self-governance have no external overseers and no central controlling servers.

The third layer in Figure \ref{fig:tech_stack_simple} transforms our usage of money.
Our proposed approach there is again the opposite of the cybercurrency community.
We believe building a trustworthy and reliable financial infrastructure from scratch is unwise.
By re-engineering and building upon existing bank-based payment systems we provide a realistic, legally compliant, and efficient overlay.
We re-use the existing infrastructure and transform it into loosely coupled autonomous entities.
We provide a proof-of-principle prototype for making international money transfers with sub-second speed, near-zero commissions, instant clearing, and real-time settlement. 
Steps towards a programmable economy necessitate a transformation of the expensive transaction network build in 1970s (SWIFT) using a programming language from the 1950s (COBOL).\footcite{scott2017society}

@@

In a recent economics publication it is stated that with blockchains, marketplaces can be bootstrapped without the need of traditional trusted intermediaries, lowering the cost of networking.\footcite{catalini2016some}
Furthermore they challenge existing revenue models and incumbents's market power, and opens opportunities for novel approaches to regulation, auctions and the provision of public goods, software, identity and reputation systems.
Little experience, evidence, and knowledge exists for devising this fourth layer.
We are currently conducting trials with various businesses to expand our understanding.
This is a new, developing area of research, however, we believe too few research teams have the resources for the learn-by-doing methodology required to make serious scientific progress.
Building open infrastructures with self-governance is costly (our key funding was the EU FP7 project P2P-Next of 19.500.000 Euro and QLectives of 6.900.000 Euro).\footcite{p2pnextfunding}\footcite{qlectivefunding}

@@

Recently the World Bank indicated that 70\% of the world’s population still lacks access to proper land titling or demarcation services.\footcite{ieglandadministration}
A digital land registration system using blockchain technology requires a re-design of existing procedures for submission and verification of records and claims.
It is most likely to demand changes to legal frameworks.

Obtaining agreement from numerous parties during several of the stages of a property transactions, detecting errors, and to preserve tamper-proof logging is a significant challenge.
Central keepers of records are often entrusted with this critical coordination task.
Blockchains clearly have potential here, offering transparency, reducing overhead, and especially providing a single-source-of-truth.
Various reports support this opinion.\footcite{deloitteblockchain}

@@

A Physical Unclonable Function (PUF) is a device that is easy and cheap to produce but practically infeasible to duplicate, due to minor variations in the manufacturing process of the hardware.\footcite{cortez2012modeling}
Their typical usage can be found in applications that require a high level of security, for instance, self-sovereign identity solutions.
The device offers tamper-proof and safe storage of cryptographic keys, representing ones identity.
Each PUF device contains an unique fingerprint, determined by randomness of embedded components.
A PUF responds to challenges and leads to unique but unpredictable responses, together forming a challenge-response pair.
These challenges are often triggered by pushing a physical button attached to the device.

@@

We present a mobile biometric-based authentication prototype that does not involve any central authority.\footcite{hammudoglu2017portable}
A proof-of-principle mobile application has been developed for Android, capable of matching fingerprints using the built-in device camera.
By only utilizing device-specific components like the embedded storage and camera, no permissioned or specialized, contractual hardware is needed.

@@

Most identity services are offered by a single authority, handing out, revoking and managing identities.
This trend continued when society adopted digital solutions: issuers of digital identities like Internet Assigned Numbers Authority (IANA) and Internet Corporation for Assigned Names and Numbers (ICANN), the organizations responsible for coordination of digital namespaces on the Internet, were and still are single, large organizations that have a large influence on the layout of the Internet landscape.\footcite{pathtossi}
Some identity schemes took a small step beyond a centralized structure and offered hierarchical solutions like certificate authorities (CAs).
However, such systems rely on root authorities which form a single point of failure.

@@

TrustChain blocks are designed to be disseminated and replicated throughout the network.
In particular, this is important when transaction history is used as input for a reputation mechanism which needs the historical behaviour of others (see Section \ref{sec:reputation}).
Additionally, replication of blocks makes the system resistant against network churn where users go on- and offline at a fast rate.
Each user operates on their own bulk storage of blocks, resulting in partial storage of the global ledger.
Collecting information of other users is challenging due to the vulnerability to various attacks, their limited resources and the burst of their interactions.
Prior work investigates this problem and propose solutions for reliable and secure collection of the interaction history.\footcite{gkoroureducing}

@@

\subsection{Scalable Consensus}
\label{sec:scalable_consensus}
Most blockchain fabrics have a requirement for a consistent state which is agreed upon by (a part of) the network.
This requirement holds in particular when considering digital money (cybercurrencies) like Bitcoin or Ethereum.
Consensus of the global ledger is necessary to prevent the double-spending attack where users transfer the same digital asset twice in multiple transactions.
Possibility of such an attack impacts reliability and trust of the overall system, preventing community adoption.
Unfortunately, most consensus mechanisms are computationally expensive and limit the global transaction throughput of the system.
For instance, the proof-of-work consensus mechanism implemented in the Bitcoin fabric limits the theoretical transaction throughput to around seven transactions per second which is by far not enough for a medium to large-sized trading platform.\footcite{duebitcoinscale}
In comparison, Visa processes several hundreds of transactions every second and in April 2017, SWIFT recorded an average of 28.38 million payments per day or around 328 per second.\footcite{visatransactions}\footcite{swifttransactions}
This motivates the need for a scalable blockchain.

@@

Our consensus mechanism proceeds in rounds and each round consists of two phases.
The outcome of each round is a set of checkpoint blocks agreed on by the facilitators of that round.
Facilitators are special user, elected during the first phase in a round of consensus.
These facilitators reach consensus not on the individual transactions like in Bitcoin or Ethereum but on the state of every chain.
If a specific user is not a facilitator, it sends its most recent checkpoint block to all facilitators.
When the facilitators received a sufficient number of checkpoint blocks, they start to reach consensus on all received checkpoint blocks using an Asynchronous Subset Consensus (ACS) algorithm.\footcite{miller2016honey}
When this algorithm is finished, the facilitators send two messages to all users, first the consensus result and second a signature message where the facilitator signed the consensus result, adding authenticity to the consensus result.
When a user receives the consensus result and a sufficient amount of valid facilitator signatures, a new checkpoint block is created and appended to the chain.
Finally, the new set of facilitators is elected and the round number is increased.
The process starts over now.

@@

\subsection{Trustworthiness and Reputation}
\label{sec:reputation}
One way to quantity trustworthiness in an economic system, is by using a reputation mechanism.
Reputation systems are widely used within the gig economy but the trust is often centred to a single authority.
Fifteen years of research on generic distributed reputation systems has yielded a wide variety of proposed designs and possible solutions.\footcite{delaviz2010improving}\footcite{delaviz2012sybilres}\footcite{kamvar2003eigentrust}
Prior work by us challenges the problem of estimating reputation using network flow algorithms.
This mechanism is called BarterCast and has been verified by a real-world implementation in our file-sharing software Tribler.\footcite{meulpolder2009bartercast}
The most challenging attack on decentralized reputation systems is the Sybil attack, where a malicious user manipulates his or her standing in the network by creating multiple fake identities (sybils) and initiating interactions with them.
To date, the Sybil attack remains largely unsolved and solutions that prevent manifestation such an attack, are often complex.

@@

Our reputation mechanism is based on PageRank, designed in 1998 by Larry Page and is called Temporal PageRank since it incorporates the notion of time.\footcite{page1999pagerank}\footcite{otte2016sybil}
PageRank is an accurate model that captures user behaviour when browsing the web and yields a ranking of websites based on relevance for that user.
The algorithm models websites and links between websites as a network and explores this network in a structured manner.
Temporal PageRank works in the same way, exploring the TrustChain data structure and analysing past transactions, including connections between them.
Additionally, it is a relatively simple and cheap technique from a computational perspective, requiring only minimal resources.
Temporal PageRank determines trustworthiness scores for other users from the perspective of a user performing the computation and is somewhat resistant against Sybil attacks in a sense that such attacks performed in the past only have minor influence on the outcome.

@@

To obtain insight in the real-world efficiency of our system, we evaluate matching efficiency of Uber taxi drives.
Uber, one of the largest companies operating in the sharing economy, can be considered as a two-sided market where taxi drivers, offering a ride to a location, are matched against users requesting transportation.
Since there is no public, reliable dataset provided by the Uber platform, we use historical information of taxi rides published by the government of New York instead.\footcite{newyorktaxi}
The dataset provides detailed temporal and geographical information of pick-up and drop-off location of each taxi ride.
We assume a total of 1100 taxi drivers and 1000 users requesting transportation, which is also close to the limitations of the hardware on which the experiment is executed.
It is also assumed that only the taxi drivers perform matching since they are likely to be connected to the network for a longer time whereas passengers most likely close the application when a taxi is available for them.
Taxi drivers and passengers are matched based on their geographical distance.
